FoRTE Disturbance Paper Analysis

### 2021-07-27 C.analysis/paper_analysis.Rmd
Update to manuscript circualted to co-authors at end of July

```{r setup, include=FALSE, message=FALSE, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width = 8, fig.height = 5)

library(dplyr)
library(ggplot2)
library(stringr)
library(stickylabeller)
library(ggpmisc)
library(ggpubr)

@@

# Define the color themes and the graphing aesthetics. 
THEME <- theme_bw()
FORTE_SEVERITY_COLORS <- c("#000000", "#009E73", "#0072B2", "#D55E00")
names(FORTE_SEVERITY_COLORS) <- c('0 %', '45 %', '65 %', '85 %')
```

@@

@@

```{r, fig.width=6.5, fig.height=4, fig.retina = 4}
ED_output %>%  
  filter(scn == "harvest_0" & variable  %in% c("GPP", "NPP", "Rh", "NEP")) %>% 
  mutate(description = paste0(variable, '\n (', unit,  ')')) %>% 
  ggplot(aes(year, value, group = met,  color = met)) + 
  geom_line() + 
  facet_wrap("description", scales = "free",  
             labeller = label_glue('({.l}) {description}')) + 
  THEME +
  labs(y = "", x = "Year")  + 
  theme(legend.position = "none") -> 
  plot

plot + 
  ggsave("figures/baseline.png", width = 6, height = 4.5)

@@

```{r}
disturbance %>%  
  mutate(Severity = severity) %>% 
  ggplot() + 
  geom_line(aes(year, value, color = Severity, line = met), alpha = 0.5) + 
  facet_wrap("variable", scales = 'free', 
             labeller = label_glue('({.l}) {variable}')) +
  THEME +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
  guides(colour = guide_legend("Severity")) + 
  labs(y = "Log Ratio", x = "Years Since Disturbance") -> 
  plot

plot
plot + 
  ggsave("figures/disturbance.png", width = 6, height = 4.5)

@@

resistance %>% 
  group_by(variable, severity) %>%  
  summarise(mean = mean(trough_resistance), sd = sd(trough_resistance))

@@

```{r, fig.width=5, fig.height=5}
ggplot(data = resistance) + 
  geom_point(aes(severity, trough_resistance, color = severity),
             position = position_jitter(), 
             alpha = 0.75, size = 1.5) + 
  facet_wrap("variable", scales = "free",  
             labeller = label_glue('({.l}) {variable}')) + 
  labs(y = "Log Ratio", 
       title = "Resistance", x = "Severity") + 
  theme_bw(base_size = 10) +
  guides(colour = guide_legend("Severity")) + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  plot

plot + 
  ggsave("figures/resistance.png", width = 6, height = 4.5)

@@

```{r}
resilience %>% 
  group_by(variable, severity) %>% 
  summarise(mean = mean(resilience))

```{r, fig.width=5, fig.height=5}
ggplot(data = resilience) + 
  geom_point(aes(severity, resilience, color = severity),              position = position_jitter(), 
             alpha = 0.75, size = 1.5) + 
  facet_wrap("variable", scales = "free",  
             labeller = label_glue('({.l}) {variable}')) + 
  labs(y = "Log Ratio", 
       title = "Resilience", 
        x = "Severity") + 
  theme_bw(base_size = 10) +
  guides(colour = guide_legend("Severity")) + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  plot

plot + 
  ggsave("figures/resilience.png", width = 6, height = 4.5)
```

@@

```{r, fig.width=5, fig.height=5}
resilience %>% 
  left_join(resistance) %>%  
  ggplot(aes(trough_resistance, resilience, color = severity)) + 
  geom_point(alpha = 0.75, size = 1.5) + 
  facet_wrap("variable", #scales = "free",  
             labeller = label_glue('({.l}) {variable}')) + 
  labs(x = "Resistance", 
       y = "Resilience", 
       title = "Resilience vs. Resistance") + 
  theme_bw(base_size = 10) +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
  guides(colour = guide_legend("Severity")) -> 
  plot

plot + 
  ggsave("figures/resistance_resilience.png", width = 6, height = 4.5)

@@

resilience %>% 
  left_join(resistance) %>%  
  ggplot(aes(trough_resistance, resilience, color = severity)) + 
  geom_point(alpha = 0.75, size = 1.5) + 
  geom_smooth(method = "lm", size = 0.5, alpha = 0.2) + 
  stat_cor(method = "pearson", aes(label = ..r.label..),
           size = 3, label.x.npc = "left", label.y.npc = "bottom", 
           r.accuracy = 0.01) + 
  labs(x = "Resistance", 
       y = "Resilience", 
       title = "Resilience vs. Resistance") + 
  theme_bw(base_size = 10) +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
    facet_wrap("variable", scales = "free",  
            labeller = label_glue('({.l}) {variable}')) + 
  guides(colour = guide_legend("Severity")) +
  ggsave("figures/resistance_resilience3.png", width = 6, height = 4.5)

@@

```{r, fig.width=10, fig.height=5.5}
monthly_met %>% 
  mutate(units = if_else(variable == "vgrd", 'm s-1', units)) %>% 
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) %>% 
  ggplot(aes(mon_time, value, color = met, group = met, line = met)) + 
  geom_line() +
   facet_wrap("description", scales = 'free', strip.position = "top",
             labeller = label_wrap_gen(width = 20)) +
  labs(y = NULL, x  = NULL) +
  THEME +
  scale_x_continuous(labels = c("Jan", "Jun", "Dec"), breaks = c(1, 6, 12)) +
  theme(legend.position = "none") -> plot

plot + 
  ggsave("figures/monthly_met.png", width = 10, height = 10)

>>>>

```{r setup, include=FALSE, message=FALSE, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width = 8, fig.height = 5)
# R 3.6.3
library(dplyr)
library(ggplot2)
library(stringr)
library(stickylabeller)
library(ggpmisc)
library(ggpubr)

@@

# Define the color themes and the graphing aesthetics. 
THEME <- theme_bw()
FORTE_SEVERITY_COLORS <- c("#009E73", "#0072B2", "#D55E00")
names(FORTE_SEVERITY_COLORS) <- c('45 %', '65 %', '85 %')
```

@@

```{r}
annual_met %>% 
  filter(variable %in% c("ugrd", "vgrd")) %>% 
  mutate(value = value^2) %>% 
  group_by(year, met) %>% 
  summarise(value = sqrt(sum(value))) %>% 
  ungroup() %>%
  mutate(units = "m s-1", 
         description = "total winnd", 
         variable = "grd") -> 
  wind_df

annual_met %>% 
  filter(!variable %in% c("ugrd", "vgrd")) %>% 
  bind_rows(wind_df) %>% 
  filter(variable != "hgt") ->
  annual_met

monthly_met %>% 
  filter(variable %in% c("ugrd", "vgrd")) %>% 
  mutate(value = value^2) %>% 
  group_by(month, mon_time, met) %>% 
  summarise(value = sqrt(sum(value))) %>% 
  ungroup() %>%
  mutate(units = "m s-1", 
         description = "total winnd", 
         variable = "grd") -> 
  wind_df

monthly_met %>% 
  filter(!variable %in% c("ugrd", "vgrd")) %>% 
  bind_rows(wind_df) %>% 
  filter(variable != "hgt") ->
  monthly_met
```

@@

```{r, fig.width=6.5, fig.height=4, fig.retina = 4}

# # UMBS forest observations taken from table 2 of 
# C.M. Gough, C.S. Vogel, H.P. Schmid, H.-B. Su, P.S. Curtis,
# Multi-year convergence of biometric and meteorological estimates of forest carbon storage,
# Agricultural and Forest Meteorology,
# Volume 148, Issue 2,
# 2008,
# Pages 158-170,
# ISSN 0168-1923,
# https://doi.org/10.1016/j.agrformet.2007.08.004
data.frame(description = c("NPP\n (MgC ha-1 year-1)", "NEP\n (MgC ha-1 year-1)"), 
           mean = c(6.54, 1.53), 
           lower =  c(6.54 - .76, 1.53 - 1.15), 
           upper = c(6.54 + .76,1.53 + 1.15 ), 
           year = c(2001, 2001)) -> 
  obs

ED_output %>%  
  filter(scn == "harvest_0" & variable  %in% c("GPP", "NPP", "Rh", "NEP")) %>% 
  mutate(description = paste0(variable, '\n (', unit,  ')')) -> 
  to_plot

  ggplot() + 
  geom_line(data = to_plot, aes(year, value, group = met), 
            alpha = 0.9, color = "#999999", ) + 
  geom_pointrange(data = obs, aes(x=year, y=mean, ymin=lower, ymax=upper)) + 
  facet_wrap("description", scales = "free", 
             labeller = label_glue('({.l}) {description}')) +
  THEME +
  labs(y = "", x = "Year")  + 
  theme(legend.position = "none") -> 
  plot

plot
ggsave(plot, filename = "figures/baseline.png", width = 6, height = 4.5)

@@

```{r}
disturbance %>%  
  mutate(Severity = severity) %>% 
  ggplot() + 
  geom_line(aes(year, value, color = Severity, line = met), alpha = 0.5) + 
  facet_wrap("variable", scales = 'free') +
            # labeller = label_glue('({.l}) {variable}')) +
  THEME +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
  guides(colour = guide_legend("Severity")) + 
  labs(y = "Log Ratio", x = "Years Since Disturbance") -> 
  plot

plot

ggsave(plot, filename = "figures/disturbance.png", width = 6, height = 4.5)

@@

resistance %>% 
  group_by(variable, severity) %>%  
  summarise(mean = mean(trough_resistance), sd = sd(trough_resistance)) %>% 
  ungroup(.) -> 
  summary_table
```

```{r}
summary_table %>% 
  select(variable, severity, mean) %>% 
  tidyr::spread(severity, mean) %>% 
  mutate(change1 = `65 %` - `45 %` / `45 %`, 
         change2 = `85 %` - `65 %` / `65 %`)

@@

```{r, fig.width=5, fig.height=5}
ggplot() + 
  geom_point(data = resistance, aes(severity, trough_resistance, color = severity),
             position = position_jitter(), 
             alpha = 0.75, size = 1.5) + 
    geom_pointrange(data = summary_table, 
                    aes(x=severity, y=mean, ymin=mean-sd, ymax=mean+sd), 
                  color = "black") + 
    ylim(-1.5, 0) +
  facet_wrap("variable",
             labeller = label_glue('({.l}) {variable}')) + 
  labs(y = "Resistance", 
       x = "Severity") + 
  theme_bw(base_size = 10) +
  theme(legend.position = "none") +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  plot

ggsave(plot, filename = "figures/resistance.png", width = 6, height = 4.5)
```

```{r}
resilience %>% 
  group_by(variable, severity) %>% 
  summarise(mean = mean(resilience), sd = sd(resilience)) %>% 
  ungroup(.) -> 
  summary_table
```

```{r}
summary_table %>% 
  select(variable, severity, mean) %>% 
  tidyr::spread(severity, mean) %>% 
  mutate(change1 = `65 %` - `45 %` / `45 %`, 
         change2 = `85 %` - `65 %` / `65 %`)

@@

```{r, fig.width=5, fig.height=5}
ggplot() + 
  geom_point(data = resilience, aes(severity, resilience, color = severity),              
             position = position_jitter(), 
             alpha = 0.75, size = 1.5) + 
  facet_wrap("variable", labeller = label_glue('({.l}) {variable}')) + 
  geom_pointrange(data = summary_table, 
                    aes(x=severity, y=mean, ymin=mean-sd, ymax=mean+sd), 
                    color = "black") + 
  labs(y = "Resilience", 
        x = "Severity") + 
  ylim(0, 0.035)+
  theme_bw(base_size = 10) +
  theme(legend.position = "none") +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  plot

  ggsave(plot, filename = "figures/resilience.png", width = 6, height = 4.5)
```

@@

```{r, fig.width=5, fig.height=5}
resilience %>% 
  left_join(resistance) %>%  
  ggplot(aes(trough_resistance, resilience, color = severity)) + 
  geom_point(alpha = 0.75, size = 1.5) + 
  facet_wrap("variable")+ #scales = "free",  
            # labeller = label_glue('({.l}) {variable}')) + 
  labs(x = "Resistance", 
       y = "Resilience", 
       title = "Resilience vs. Resistance") + 
  theme_bw(base_size = 10) +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
  guides(colour = guide_legend("Severity")) -> 
  plot

ggsave(plot, filename = "figures/resistance_resilience.png", width = 6, height = 4.5)

@@

resilience %>% 
  left_join(resistance) %>%  
  ggplot(aes(trough_resistance, resilience, color = severity)) + 
  geom_point(alpha = 0.75, size = 1.5) + 
  geom_smooth(method = "lm", size = 0.5, alpha = 0.2) + 
  stat_cor(method = "pearson", 
           aes(label = ..r.label..),
           size = 3, 
           label.y.npc = "bottom",
           r.accuracy = 0.01) + 
  labs(x = "Resistance", 
       y = "Resilience", 
       title = "Resilience vs. Resistance") + 
  theme_bw(base_size = 10) +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
    facet_wrap("variable", 
               scales = "free", 
               labeller = label_glue('({.l}) {variable}')) + 
  guides(colour = guide_legend("Severity")) -> 
  plot
  ggsave(plot, filename = "figures/resistance_resilience3.png", width = 6, height = 4.5)

@@

```{r, fig.width=10, fig.height=5.5}
monthly_met %>% 
  mutate(units = if_else(variable == "vgrd", 'm s-1', units)) %>% 
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) -> 
  data

data %>% 
  filter(met == "2019") -> 
  data_2019

ggplot() + 
  geom_line(data = data, aes(mon_time, value, group = met, line = met), 
            alpha = 0.8, color = "#999999") +
  geom_line(data = data_2019, aes(mon_time, value, color = met, group = met, line = met), 
            color = "#D55E00", size = 1) +
   facet_wrap("description", scales = 'free', strip.position = "top",
             labeller = label_wrap_gen(width = 20), ncol = 2) +
  labs(y = NULL, x  = NULL) +
  THEME +
  scale_x_continuous(labels = c("Jan", "Jun", "Dec"), breaks = c(1, 6, 12)) +
  theme(legend.position = "none") -> 
  plot

plot

ggsave(plot, filename = "figures/monthly_met.png", width = 6, height = 10)
```
